Evaluating Cultural Transmission in Bronze Age burial rites of Central, Northern and North-western Europe using radiocarbon data

### 2019-05-23 review/answer_to_review_2.md

> Page 2, Column 1, Line 29. ‘simplify survival’ → increase?

Rephrased.

> Page 2, Column 1, Line 42. ‘seemlingly random’  This is one of those tricky phrasing that has different meaning depending on the readership. ‘Random’ to me almost equals to unbiased transmission with drift, and stating that it is ‘seemlingly’ reads as if in reality this is not the case. This is perfectly fine, but it is slightly at odds with the definition of ‘style’ and the implicit assumption of neutrality. I would consider rephrasing, just to avoid potential misunderstandings.

You're right. I thought a lot about this and finally decided to simply remove the term "seemlingly random". In my opinion the binary distinction between style and function reaches its limit with burial traditions. In many cultural contexts, death and funerals are strongly reflected. Much more thought is given to this than to patterns on pots. The simulation paper will have to look at this difference more carefully, because there the neutrality is an even more important assumption.

@@

> Page 3, Column 2, Line 59. ‘the region are circular’ Why this shape if it causes overlap? Why not hexagons?

This is true. I added a sentence to point this out. For this paper I will use the established regions, because changing them at this point would render the area discussions useless. In the future more data might be available and it should be possible to establish a lot more fine grained, hexagonal or rectangular region grid.

> Page 5, Figure 3. Just an idea for a follow-up study. Given these are effectively adoption curves, one could fit a logistic model to assess differences in the rate of adoption and estimates on the timing of earliest uptake. It would be also very interesting to determine whether indeed there is a long-tail in southern Germany and what this means (cf. Henrich 2001 on American Anthropologist, but see also recent debates on adoption curves, e.g. https://www.biorxiv.org/content/biorxiv/early/2017/11/16/159038.full.pdf). Food for thoughts.

Well noted. Thank you! Incidentally for Southern Germany there is a paper by Frank Falkenstein that reconstructed the adoption rate of these burial traditions based on archaeologically dated graves:

Falkenstein, F. (2017). Zum Wandel der Bestattungssitten von der Hügelgräber- zur Urnenfelderkultur in Süddeutschland. In Brandherm, D. and Nessel, B., editors, Phasenübergänge Und Umbrüche Im Bronzezeitlichen
Europa, number 297 in UPA, pages 77–96. Bonn.
https://www.academia.edu/36327050/Zum_Wandel_der_Bestattungssitten_von_der_H%C3%BCgelgr%C3%A4ber-_zur_Urnenfelderkultur_in_S%C3%BCddeutschland

### 2019-05-23 review/answer_to_review_2.md

Fixed.

# Reviewer: 2

Comments to the Author:

> I read the revised version of the manuscript and I think the authors have addressed most of the issues I raised, and acknowledged others in the text. I am still unconvinced by the creation of the artificial subsets and think that perhaps using temporal distance matrix (see for example Shennan et al 2015 on Evolution and Human Behaviour) might have been a solution, but I also see the point that such approach would not directly explore time-series. Ultimately this is an instance of the so-called MAUP (Modifiable Areal Unit Problem) in geography, and perhaps there are some more nuanced approaches out there. I think for the purpose of this paper acknowledging the problem is more than sufficient.

Thank you very much for your review. You gave me some great ideas how this problem could be approached in a better way and your review will be the first thing I will go to, when I have the opportunity to continue the work on this front. For measuring the degree of interaction in a better way and withput regions, I recently got some inspiration from a paper by Loog et al 2017:

Loog, L., Lahr, M. M., Kovacevic, M., Manica, A., Eriksson, A., & Thomas, M. G. (2017). Estimating mobility using sparse data: Application to human genetic variation. Proceedings of the National Academy of Sciences, 114(46), 12213–12218. https://doi.org/10.1073/pnas.1703642114

Two major take aways: (1) Their mobility estimator could be used for a better, region independent statistic for the general degree of interaction. (2) I could as well use a moving window instead of 200 years time slices (Figure 6).

I added some sentences in the conclusion to give an outlook on future work.

> I also still have some issues with how sample size is effectively being ignored once proportions are computed, and I think this should be emphasised more strongly as I expect not everyone would be reading beyond the conclusion section.

Added sentence to the conclusion to make this very clear.

### 2019-05-23 review/answer_to_review_2.md

# Reviewer: 2

Comments to the Author:

> I read the revised version of the manuscript and I think the authors have addressed most of the issues I raised, and acknowledged others in the text. I am still unconvinced by the creation of the artificial subsets and think that perhaps using temporal distance matrix (see for example Shennan et al 2015 on Evolution and Human Behaviour) might have been a solution, but I also see the point that such approach would not directly explore time-series. Ultimately this is an instance of the so-called MAUP (Modifiable Areal Unit Problem) in geography, and perhaps there are some more nuanced approaches out there. I think for the purpose of this paper acknowledging the problem is more than sufficient.

Thank you very much for your review. You gave me some great ideas how this problem could be approached in a better way and your review will be the first thing I will go to, when I have the opportunity to continue the work on this front. For measuring the degree of interaction in a better way and without artificial regions, I recently got some inspiration from a paper by Loog et al 2017:

### 2019-05-23 article/04_Discussion.Rmd
the negative correlation is not significant

Against this background, the observation that there is no permanent, significant correlation either between the variables or with the spatial distance is all the more astonishing (Table \ref{tab:dacta} and \ref{tab:dactb}, Figure \ref{correlationtimeseries}). Only for certain periods of time this correlation is visible in the Radon-B data and the *burial construction* distance shows a higher degree of correlation with spatial distance than the *burial type* distance, which even seems to behave in an opposing manner: negative correlation. If one assumes the reconstructed timelines of cultural similarity intensity to be correct, than one has to explain a rise (between 2200 and 2000 calBC) and a decline (after 2000 calBC) of cultural-spatial-distance correlation. There are several reasons why any interpretation of this observation is difficult so far:

>>>>

Against this background, the observation that there is no permanent, significant positive correlation either between the variables or with spatial distance is all the more astonishing (Table \ref{tab:dacta} and \ref{tab:dactb}, Figure \ref{correlationtimeseries}). Only for certain periods of time this correlation is visible in the Radon-B data. The *burial construction* distance generally shows a higher degree of correlation with spatial distance than the *burial type* distance, which even seems to behave in an opposing manner: negative correlation (also not statistically significant!). If one assumes the reconstructed timelines of cultural similarity intensity to be correct, than one has to explain a rise (between 2200 and 2000 calBC) and a decline (after 2000 calBC) of cultural-spatial-distance correlation. There are several reasons why any interpretation of this observation is difficult so far:
